Highlights

### 2018-01-22 paper/final/Highlights.tex
review of Sec 6 and remark added in Sec3

\begin{itemize}
	% Remember maximum 85 characters (included spaces, points, commas, etc.)
	\item Predictive control strategy using data-based models via machine learning algorithms.
	\item Data Predictive Control shows comparable performance to Model Predictive Control.
	\item Data Predictive Control shows good performance in Demand Response curtailment.
	\item Data Predictive Control allows energy saving up to $50\%$ in the considered case study.
\end{itemize}

>>>>

\begin{itemize}
	% Remember maximum 85 characters (included spaces, points, commas, etc.)
	\item Predictive control strategy using data-driven models via Random Forests.
	\item Data Predictive Control shows comparable performance to Model Predictive Control.
	\item Data Predictive Control shows good performance in Demand Response curtailment.
	\item Data Predictive Control allows energy saving up to $50\%$ in considered case study.
	\item Data Predictive Control is robust to data and weather forecast uncertainties.
\end{itemize}

### 2018-01-22 paper/final/casestudyReal.tex
review of Sec 6 and remark added in Sec3

\section{CASE STUDY: OPTIMAL HEATING SYSTEM SCHEDULING}
\label{S:realCaseStudy}

In this section, we show an application of DPC to a real house located in L'Aquila, Italy. Random forest models for DPC built using historical data from this house are described in Section \ref{SS:randomForestsModels}. 
In Section \ref{SS:energyPlusmodel}, we present the thermal energy model of the building in EnergyPlus, built using historical data, construction layout and materials after spending $\sim3$ months of efforts.
In Section \ref{SS:controllersDPCandBangbang}, we set up the DPC optimization problem and describe the bang-bang control strategy.
Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy savings using the EnergyPlus model.
\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC provides up to $49.2\%$ energy savings with respect to the bang-bang controller while guaranteeing thermal comfort for the occupants, considering the perfect knowledge of the weather forecast. 
In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
The section organization is graphically shown in Figure \ref{F:overview}.
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller. DPC optimzation uses Random Forests with affine functions as predictive models. Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}

\textcolor[rgb]{0,0,1}{The contribution of this section is threefold.

@@

	\item We consider experimental data from an occupied house, hence subject to real world imperfections, such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc.
	This is different from Section \ref{S:casestudy} where we used data simulated from an EnergyPlus model.
	\item We modify the DPC algorithm to take into account uncertainties In the weather forecast, making DPC robust with respect to prediction errors.

@@

@@

\textcolor[rgb]{0,0,1}{We considered two different simulation conditions to show the robustness of our approach.
\begin{enumerate}
	\item In Section \ref{SSS:DisturbancePerfect}, we provide the results considering perfect knowledge of the weather forecast.
	\item In Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainties. Hence, we modify Problem \ref{P:dpcRealCase} by adding a Gaussian noise with high variance to the perfect forecast, and show that the results are close to the perfect forecast case.
	This verifies our methodology is robust with respect to disturbance uncertainties.
\end{enumerate}
\subsubsection{Perfect knowledge of the weather forecast}\label{SSS:DisturbancePerfect}
In this section, we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
\paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}. For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from $00$ hrs on May $1$, $2016$ to  $13$ hrs on May $5$, $2016$. \textcolor[rgb]{0,0,1}{The whole period will be used in ``Result 2" to provide the bounds violation errors, and in ``Result 3" for the energy consumption comparison.}
We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$. Bang-bang control also presents small bounds violations due to its working principle. We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range. 
We will see in Figure \ref{F:comparison_all_energy_E+} that this translates to significant energy saving.

@@

\textcolor[rgb]{0,0,1}{
\subsubsection{Weather forecast subject to uncertainty}\label{SSS:DisturbanceUncertain}
The authors in \cite{Petersen2014AE} investigate the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compare against a rule-based strategy.
They consider 48 different scenarios of uncertainties for 72 hours weather forecast.
With such a long horizon, results have shown that (with a few exceptions) MPC outperforms the rule-based controller in providing energy savings, and is in general quite close to the perfect forecast case despite the uncertainty in weather forecast.}

\textcolor[rgb]{0,0,1}{The length of horizon for the DPC algorithm presented in this paper is usually much shorter, for example $6$ hours in Section \ref{S:proof}, $7$ hours in Section \ref{S:casestudy}, and $40$ minutes in Section \ref{S:realCaseStudy}. 
%This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
We next show the robustness of our approach considering noisy weather forecast.}

\textcolor[rgb]{0,0,1}{We add noise with normal distribution to the disturbance variables in $\X^d$ which is used to obtain parameters $\hat \Theta$ in the DPC problem \eqref{E:DPCrealcase} i.e.~the disturbance in the input in the blue rectangle in Figure \ref{F:overview}, while we use the true values to simulate the process i.e.~the disturbance in the input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}.
We summarize the mean and the standard deviation of the Gaussian noises that we add to each variable in Table \ref{T:NoiseParameters}.
Since ``Time of the day" and ``Day of the week" are perfectly known, we do not add any noise to these variables.
We assume the noise to have zero mean.
Table \ref{T:NoiseParameters} also lists the range of the variables in the historical dataset.
This is to show that the standard deviation we choose adds significant error to the forecast.
For example, a standard deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$.
We solve the DPC problem \ref{P:dpcRealCase} considering the $\hat \Theta$ obtained using the noisy forecast.
We obtain the following results. We show the change in the performance on bounds violation and energy consumption due to inaccurate forecast.}
\begin{table}[t!]
	\centering	
	\textcolor[rgb]{0,0,1}{\begin{tabular}{lccc}
		\toprule
		Variable               & Range      & Mean & Deviation \\ 
		\midrule
		Outside temperature    & [0,31]     & 0	   & 0.5       \\
		Wind                   & [0,5] 		& 0    & 0.25      \\
		Atmospheric pressure   & [990,1030] & 0    & 50        \\
		Relative Humidity      & [20,91]	& 0    & 5         \\
		Solar Radiation        & [0,1000]   & 0    & 50        \\
		Time of the day        & [0,23]     & 0    & 0         \\
		Day of the week        & [1,7]      & 0    & 0         \\
		\bottomrule
	\end{tabular}}
	\caption{\textcolor[rgb]{0,0,1}{Mean and standard deviation of the Gaussian noise added on the weather forecast data, and the range of the variables in the historical dataset.}}
	\captionsetup{justification=centering}
	\label{T:NoiseParameters}
\end{table}

\textcolor[rgb]{0,0,1}{\paragraph{Result 1} Figure \ref{F:ComparisonTempNoisy} shows a comparison of the DPC results in terms of temperature control for thermal comfort between the perfect and the noisy forecast.
For the sake of clarity, we show only \(4\frac{1}{2}\) days of the 15 days simulation period. We split the plot into 3 sub-figures referring to the small, medium and large violation cases.
We see that, except for an isolated case in the medium violation case (slightly before ``02/05"), we have a small performance deterioration in terms of bounds violation. The effect is more prominent in the large violation case.
However, the maximum violation is still lower than $1\,\degree C$, as in the case of perfect forecast.
We report the violation errors in terms of MBE and CV(RMSE) (in $\%$) computed for the whole simulative period (15 days) in Table \ref{T:violationErrorsNoisy}.
Comparing them with the values obtained earlier in Table \ref{T:violationErrors}, we see that the performance of DPC is robust to noisy predictions despite a large error in the weather forecast.
}
\begin{figure}[t!]
	\begin{center}
		\vspace{1.1cm}
		\subfigure{
			\label{F:SmallNoisy}
			\centering
			\includegraphics[width=23pc]{figures/Temperatures_small_noisy.eps}
		}	
		\subfigure{
			\label{F:MediumNoisy}
			\centering
			\includegraphics[width=23pc]{figures/Temperatures_medium_noisy.eps}
		}
		\subfigure{
			\label{F:LargeNoisy}
			\centering
			\includegraphics[width=23pc]{figures/Temperatures_large_noisy.eps}
		}	
	\end{center}
	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC considering perfect and noisy weather forecast. Small, medium and large violation cases are compared.}}
	\captionsetup{justification=centering}
	\label{F:ComparisonTempNoisy}
\end{figure}

@@

\textcolor[rgb]{0,0,1}{\paragraph{Result 2} 
In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (solid line) and the noisy forecast (dashed line).
We observe that the energy usage is very close in both cases.
Recall, we optimize a weighted sum of energy and thermal comfort, so depending on the conditions for which the disturbance is over or underestimated, the MPC solution changes depending upon the prediction errors.
In the small violation case, the energy consumption of the two conditions is closer than in the others.
This is because the DPC keeps the temperature within the bounds with an extremely small error in bounds violation.
In particular, the only exception can be seen before May $3^{rd}$ (03/05), where, due to the prediction error, the temperature violates the bounds, so the control uses more energy (than the perfect case) to bring it back above the threshold.
In the other two cases, when the temperature violates the bounds, due to the fact that violation constraints are more relaxed, the DPC does not require much larger control effort, hence using lesser energy.}

>>>>

\section{APPLICATION TO OPTIMAL HEATING SYSTEM SCHEDULING}
\label{S:realCaseStudy}

In this section we show an application of DPC to a real house located in L'Aquila, Italy. Random forest models for DPC built using historical data from this house are described in Section \ref{SS:randomForestsModels}. 
In Section \ref{SS:energyPlusmodel}, we present the thermal energy model of the building in EnergyPlus, built using historical data, construction layout and materials after spending $\sim3$ months of efforts.
In Section \ref{SS:controllersDPCandBangbang}, we set up the DPC optimization problem and describe the bang-bang control strategy.
Finally, in Section \ref{SS:simulationResults}, the performance of the two controllers are compared in terms of energy savings using the EnergyPlus model.
\textcolor[rgb]{0,0,1}{In particular, in Section \ref{SSS:DisturbancePerfect}, we show that DPC provides up to $49.2\%$ energy savings with respect to the bang-bang controller while guaranteeing thermal comfort for the occupants, considering the perfect knowledge of the weather forecast. 
In Section \ref{SSS:DisturbanceUncertain}, we show that DPC is robust with respect to imperfect weather forecast.}
\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\linewidth]{figures/overview.eps}
		\caption{Random Forest trained on all features including the control variable are used as the plant in the closed-loop simulations with DPC and bang-bang controller. DPC optimization uses Random Forests with affine functions as predictive models. Controllers' strategies are evaluated using the EnergyPlus model to compare the actual energy consumption.}
		\captionsetup{justification=centering}
		\label{F:overview}
	\end{center}
\end{figure}

\textcolor[rgb]{0,0,1}{The contribution of this section, graphically shown in Figure \ref{F:overview}, is threefold.

@@

	\item Differently from Section \ref{S:casestudy}, where we used data simulated from an EnergyPlus model, we consider here real data from an occupied house, hence subject to real world imperfections such as random (non predictable) occupancy schedules, open/close windows, random (non predictable) light on/off switch, etc.

@@

	\item We consider uncertainties in the weather forecast, showing that DPC is robust with respect to prediction inaccuracies.

@@

\textcolor[rgb]{0,0,1}{We considered two different simulation conditions to show the robustness of our approach.
\begin{enumerate}
	\item In Section \ref{SSS:DisturbancePerfect}, we provide the results considering perfect knowledge of the weather forecast and only focus on robustness with respect to uncertainties due to real data acquisition.
	\item In Section \ref{SSS:DisturbanceUncertain}, we consider weather forecast subject to uncertainties. We modify Problem \ref{P:dpcRealCase} by adding Gaussian noise to the perfect forecast and show that the results are close to the perfect weather forecast case.
\end{enumerate}
\subsubsection{Perfect knowledge of the weather forecast}\label{SSS:DisturbancePerfect}
In this section, we ran the simulations considering perfect knowledge of the disturbance over the horizon, obtaining the following results.}
\paragraph{Result 1} The comparison for temperature and input schedule obtained allowing small bound violations in DPC is shown in Figure \ref{F:comparison_small}. For sake of the plot's clarity, the shown period is restricted to $4$ days and a half, from $00$ hrs on May $1$, $2016$ to  $13$ hrs on May $5$, $2016$. \textcolor[rgb]{0,0,1}{The whole period will be used in ``Result 2" to provide the bounds violation errors, and in ``Result 3" for the energy consumption comparison.}
We can see that the temperature controlled with DPC does not violate the bounds and if it does then the violation is approximately $0.1\,\degree C$. Bang-bang control also presents small bounds violations due to its working principle. We can see how the DPC control law requires the heating system to be ON for less time than the bang-bang one to keep the temperature in the comfort range. 
We will see in Figure \ref{F:comparison_all_energy_E+} that this translates to significant energy saving.

@@

\subsubsection{Weather forecast subject to uncertainty}\label{SSS:DisturbanceUncertain}

\textcolor[rgb]{0,0,1}{In this section we study the robustness of our approach to noisy weather forecast. To this aim, we add noise with a zero-mean Gaussian distribution to the prediction variables $\tilde d(k+1),\ldots,\tilde d(k+N)$ used to obtain the parameters $\hat \Theta_{\cdot_j}$, $j=0,\ldots,N$ in the DPC Problem \eqref{E:DPCrealcase} (i.e.~the disturbance in the input in the blue rectangle in Figure \ref{F:overview}) while we use the true values to simulate the process (i.e.~the disturbance in the input in the box ``Plant" in the red rectangle in Figure \ref{F:overview}).\\
We summarize the mean and the standard deviation of the Gaussian noises that we add to each variable in Table \ref{T:NoiseParameters}. For example, a standard deviation of $0.5\degree C$ on the outside air temperature means that the error on the predicted temperature lies within the range of $\pm 1.5\degree C$ with a probability of $99\%$. Since ``Time of the day" and ``Day of the week" are perfectly known, we clearly do not add any noise to these variables. Table \ref{T:NoiseParameters} also provides the range of values of the variables in the historical disturbances dataset to emphasise that the standard deviation we choose adds significant error to the forecast.
}

\begin{table}[t!]
	\centering	
	\textcolor[rgb]{0,0,1}{\begin{tabular}{lccc}
		\toprule
		Variable               & Range      & Mean & Deviation \\ 
		\midrule
		Outside temperature    & [0,31]     & 0	   & 0.5       \\
		Wind                   & [0,5] 		& 0    & 0.25      \\
		Atmospheric pressure   & [990,1030] & 0    & 50        \\
		Relative Humidity      & [20,91]	& 0    & 5         \\
		Solar Radiation        & [0,1000]   & 0    & 50        \\
		Time of the day        & [0,23]     & 0    & 0         \\
		Day of the week        & [1,7]      & 0    & 0         \\
		\bottomrule
	\end{tabular}}
	\caption{\textcolor[rgb]{0,0,1}{Mean and standard deviation of the Gaussian noise added on the weather forecast data, and the range of the variables in the historical dataset.}}
	\captionsetup{justification=centering}
	\label{T:NoiseParameters}
\end{table}

\textcolor[rgb]{0,0,1}{The following paragraphs provide comments on the change in performance due to inaccurate forecast on bounds violation and on energy consumption.\\
\paragraph{Result 1} Figure \ref{F:ComparisonTempNoisy} shows a comparison of the DPC results in terms of temperature control for thermal comfort between the perfect and the noisy forecast.
For the sake of clarity, we show only \(4\frac{1}{2}\) days of the 15 days simulation period. We split the plot into 3 sub-figures referring to the small, medium and large violation cases.
We see that, except for an isolated case in the medium violation case (slightly before ``02/05"), we have a small performance deterioration in terms of bounds violation. The effect is more prominent in the large violation case.
However, the maximum violation is still lower than $1\,\degree C$, as in the case of perfect forecast.
We report the violation errors in terms of MBE and CV(RMSE) (in $\%$) computed for the whole simulative period (15 days) in Table \ref{T:violationErrorsNoisy}.
Comparing them with the values obtained earlier in Table \ref{T:violationErrors}, we see that the performance of DPC is robust to noisy predictions despite a large error in the weather forecast.
}
\begin{figure}[t!]
	\begin{center}
		\vspace{1.1cm}
		\subfigure{
			\label{F:SmallNoisy}
			\centering
			\includegraphics[width=23pc]{figures/Temperatures_small_noisy.eps}
		}	
		\subfigure{
			\label{F:MediumNoisy}
			\centering
			\includegraphics[width=23pc]{figures/Temperatures_medium_noisy.eps}
		}
		\subfigure{
			\label{F:LargeNoisy}
			\centering
			\includegraphics[width=23pc]{figures/Temperatures_large_noisy.eps}
		}	
	\end{center}
	\caption{\textcolor[rgb]{0,0,1}{Comparison of DPC considering perfect and noisy weather forecast. Small, medium and large violation cases are compared.}}
	\captionsetup{justification=centering}
	\label{F:ComparisonTempNoisy}
\end{figure}

@@

\textcolor[rgb]{0,0,1}{\paragraph{Result 2} 
In Figure \ref{F:comparison_all_energy_E+_noisy}, we show the difference in terms of energy consumption considering the perfect weather forecast (solid line) and the noisy forecast (dashed line). We first recall that our optimization function is a weighted sum of energy and thermal comfort: when the weather forecast is affect to inaccuracies the MPC solution (and thus the control input) changes depending upon the prediction errors.\\
We immediately observe that the energy usage in the ideal and perturbed cases is very close, especially in the small violation case. Our interpretation for this is that the DPC, in the small violation case, keeps the temperature within the bounds with an extremely small error in bounds violation. The only exception can be seen before May $3^{rd}$ (03/05) where, due to a large prediction error, the temperature violates the bounds: as a consequence the controller uses much more energy (with respect to the perfect forecast knowledge case) to bring the temperature in the bounds, as evidenced by the blacked lines of Figure \ref{F:comparison_all_energy_E+_noisy}. In the medium and large violation cases, instead, when the temperature violates the bounds the DPC does not strongly increase the energy usage as the violation tolerance is more relaxed: this produces a general reduction of energy usage in the "noisy" simulations, as evidenced by the purple and red lines of Figure \ref{F:comparison_all_energy_E+_noisy}.
}

### 2018-01-22 paper/final/dpc.tex
review of Sec 6 and remark added in Sec3

\section{DATA PREDICTIVE CONTROL}

\label{S:dpc}

@@

\subsection{DPC-RT: DPC with Regression Trees}

\label{SS:dpcrt}

@@

\begin{figure}[t!]

	\centering

	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}

	\caption{\textit{Step 1:} Tree $\mathcal{T}_j$ trained only with variables in $\X^d$ using adapted RT algorithm. Tree $\mathcal{T}'_j$ trained with variables in $\X^d$ and control variables in $\X^c$ using classical RT algorithm. \textit{Step 2:} In the leaf $\ell_i$ of the tree an affine model $\beta_{ij}$ is defined as a function only of the control variables dataset corresponding to leaf $\ell_i$.}

	\captionsetup{justification=centering}

	\label{F:dpc-sepvars}

\end{figure}

@@

\textcolor[rgb]{0,0,1}{The next problem now is: \emph{how to use this modeling framework to set up an MPC problem?} We setup the MPC optimization problem in the general case of multiple responses, i.e. $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, as follows:}

\begin{problem}\label{P:dpcrt}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}     =   \hat \beta_{j} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                       \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                                   \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										              \\
		&                                       & & \epsilon_j \geq  0							                                                  \\
		&                                       & & j           =    0,\ldots,N,               									                  \\
		\end{aligned}
		\label{E:dpcrt}
	\end{equation}
\end{problem}

@@

@@

\subsection{DPC-En: DPC with Ensemble Methods}

\label{SS:dpcrf}

@@

%In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.

%Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.

\textcolor[rgb]{0,0,1}{We can now set up the MPC problem for DPC-En as follows:}
\begin{problem}\label{P:dpcrf}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{j} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
		&                                       & & \epsilon_j \geq  0							                                     \\
		&                                       & & j           =    0,\ldots,N,            									         \\
		\end{aligned}
		\label{E:dpcrf}
	\end{equation}
\end{problem}
\noindent \textcolor[rgb]{0,0,1}{where $\hat{\Theta}_{j}$ is defined later on. Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$. We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$. In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{j} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$. Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{j}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section. The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}

\begin{figure}[t!]
	\centering
	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
	\caption{\textcolor[rgb]{0,0,1}{Graphical description of the procedure for solving Problem \ref{P:dpcrf} at each time $k$}}
	\label{F:dpc-algo-rf}
\end{figure}

>>>>

\section{DATA PREDICTIVE CONTROL}
\label{S:dpc}

@@

\subsection{DPC-RT: DPC with Regression Trees}
\label{SS:dpcrt}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
	\caption{\textit{Step 1:} Tree $\mathcal{T}_j$ trained only with variables in $\X^d$ using adapted RT algorithm. Tree $\mathcal{T}'_j$ trained with variables in $\X^d$ and control variables in $\X^c$ using classical RT algorithm. \textit{Step 2:} In the leaf $\ell_i$ of the tree an affine model $\beta_{ij}$ is defined as a function only of the control variables dataset corresponding to leaf $\ell_i$.}
	\captionsetup{justification=centering}
	\label{F:dpc-sepvars}
\end{figure}

@@

\textcolor[rgb]{0,0,1}{The next problem now is: \emph{how to use this modeling framework to set up an MPC problem?} We setup the MPC optimization problem in the general case of multiple responses, i.e. $y(k)\in\mathbb{R}^{n_y},\ n_y\geq 1$, as follows:}
\begin{problem}\label{P:dpcrt}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}     =   \hat \beta_{j} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                       \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                                   \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										              \\
		&                                       & & \epsilon_j \geq  0							                                                  \\
		&                                       & & j           =    0,\ldots,N,               									                  \\
		\end{aligned}
		\label{E:dpcrt}
	\end{equation}
\end{problem}

@@

\textcolor[rgb]{0,0,1}{\begin{remark}
The authors in \cite{Petersen2014AE} investigate the effect of uncertainty in the weather forecast on the performance of MPC in building systems operations through a large-scale simulation study, and compare against a rule-based strategy. They consider 48 different scenarios of uncertainties for 72 hours weather forecast. With such a long horizon, results have shown that (with a few exceptions) MPC outperforms the rule-based controller in providing energy savings, and is in general quite close to the perfect forecast case despite the uncertainty in weather forecast. The length of horizon for the DPC algorithm presented in this paper is usually much shorter, for example $6$ hours in Section \ref{S:proof}, $7$ hours in Section \ref{S:casestudy}, and $40$ minutes in Section \ref{S:realCaseStudy}. As a consequence, we can reasonably presume that our approach will be robust to weather forecast inaccuracies. Indeed, in Section \ref{S:realCaseStudy} we test the robustness of our approach to noisy weather forecast and show that the control performance is very close to the ideal case.
%This means the weather forecast is more accurate than the 72 hours case presented in \cite{Petersen2014AE}, hence also control performance is even closer to the perfect knowledge forecast case.
\end{remark}}

@@

\subsection{DPC-En: DPC with Ensemble Methods}
\label{SS:dpcrf}

@@

%In particular, as mentioned above, each tree $\T_i$ of each forest $\F_j$ is trained using $(\bar \X^d_i,\bar\Y_i)$.
%Here $(\bar\X^c_i,\bar\Y_i)$ correspond to the in-bag samples for the trees.

\textcolor[rgb]{0,0,1}{We can now set up the MPC problem for DPC-En as follows:}
\begin{problem}\label{P:dpcrf}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{j} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
		&                                       & & \epsilon_j \geq  0							                                     \\
		&                                       & & j           =    0,\ldots,N,            									         \\
		\end{aligned}
		\label{E:dpcrf}
	\end{equation}
\end{problem}

\noindent \textcolor[rgb]{0,0,1}{where $\hat{\Theta}_{j}$ is defined later on. Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$. We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$. In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{j} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$. Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{j}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section. The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}
\begin{figure}[t!]
	\centering
	\includegraphics[width=26pc]{figures/dpc-algo-rf.eps}
	\caption{\textcolor[rgb]{0,0,1}{Graphical description of the procedure for solving Problem \ref{P:dpcrf} at each time $k$}}
	\label{F:dpc-algo-rf}
\end{figure}
